A Formal Model of Checked C

### 2023-01-17 sandp23/background.tex

\liyi{why we need the following? I suggest we can cut. }
Oftentimes, inserted checks can be optimized away by LLVM.
Consider the above code to be enclosed in another condition, such as,~\inlinecode{if (n<u) if (n>0) a[n-1] =}.
In such cases, the inserted check can be
removed as the outer condition $\inlinecode{n<u}$ already ensures that the
dereference is within bounds.
Dual~\etal~\cite{duanrefactoring} found essentially no
overhead when running Checked C-converted portions of the FreeBSD kernel.

>>>>

\liyi{why we need the following? I suggest we can cut. }
Oftentimes, inserted checks can be optimized away by LLVM.
For instance, Dual~\etal~\cite{duanrefactoring} found essentially no
overhead when running Checked C-converted portions of the FreeBSD kernel.
%Consider the above code to be enclosed in another condition, such as,~\inlinecode{if (n<u) if (n>0) a[n-1] =}.
%In such cases, the inserted check can be
%removed as the outer condition $\inlinecode{n<u}$ already ensures that the
%dereference is within bounds.
%Dual~\etal~\cite{duanrefactoring} found essentially no
%overhead when running Checked C-converted portions of the FreeBSD kernel.

### 2023-01-17 sandp23/implementation.tex

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions without memory modification allowing developers to use the function freely in ~\cregion.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the above _TLIB annotation.
strncat(dst, t_str, n);
\end{minted}

Passing tainted pointer ~\code{t_str} to unqualified ~\code{strncat} above is disallowed by our type-checker as it expects both the pointers to be non-tainted since we are in ~\cregion. If the developer knows that~\code{strncat} accesses only~\code{n} bytes from the source buffer, she would like to directly use ~\code{strncat} with tainted source pointer in ~\cregion, instead of implementing the logic for string concatenation in ~\ucregion or marshalling data between the two regions.  
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
- extern char *strncat (char *__restrict __dest,
+ _TLIB extern char *strncat (char *__restrict __dest,
const char *__restrict __src, size_t __n); // In the header file
}
\end{minted}

Consequently, annotating ~\code{strncat} as shown above relaxes type-checking for all the arguments to its calls. However, \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying, thereby, requiring developers to be careful to while using ~\code{\_TLIB} to prevent memory address leaks.

>>>>

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions without memory modification allowing developers to use the function freely in ~\cregion.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the _TLIB annotation below.
strncat(dst, t_str, n);
\end{minted}

\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{diff}
- extern char *strncat (char *__restrict __dest,
+ _TLIB extern char *strncat (char *__restrict __dest,
const char *__restrict __src, size_t __n); // In the header file
}
\end{minted}

Passing tainted pointer \code{t_str} to unqualified ~\code{strncat} above is disallowed without having additional sandboxing for \code{t_str}. If a user ascertains that \code{t_str} has the right buffer size for \code{strncat}, she might label \code{strncat} with \code{\_TLIB}, so that \code{t_str} is treated as an unchecked pointer; such annotation relaxes type-checking for all the arguments to its calls. It is worth noting that \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying; therefore, using \code{\_TLIB} requires users' awareness of memory address leaks.
